---
title: "Online Shoppers Purchasing Intention - Analysis"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: show
execute:
  warning: false
  message: false
  author: Arttu Isosaari
---

## Introduction

This analysis examines customer shopping behavior using the Online Shoppers Purchasing Intention dataset. The dataset is open data and contains observations describing individual website visits with information such as timing, visitor type, and traffic source. The response variable is Revenue (TRUE/FALSE), which indicates whether a visit leads to a purchase.

The analysis focuses on three research questions:

1.  Are people more likely to buy on weekends or weekdays?
2.  Do returning customers have a better conversion rate than new customers?
3.  Does purchasing depend on the channel the customer came from?

The main method is the χ² test, which only tells us about possible dependence between variables. It does not reveal causality, so we interpret the results with, for example, probability tables.

```{r}
#| label: load-data
#| include: false

# load data
OSPI <- read.csv("online_shoppers_intention.csv")

# Inspect data structure
str(OSPI)
head(OSPI)
```

## Analysis 1: Weekend vs. Weekday

We examine whether purchase likelihood differs between weekends and weekdays.

```{r}
#| label: analysis-weekend
#| echo: true


# Run chi-square test
tab1 <- table(OSPI$Weekend, OSPI$Revenue)
testi1 <- chisq.test(tab1)
tulokset1 <- c(testi1$statistic, testi1$p.value)
names(tulokset1) <- c("Chi-square", "p-value")
tulokset1
```

We reject the null hypothesis at all commonly used significance levels, indicating a statistical relationship between purchases and weekends.

Let's look at this more closely by creating a probability table:

```{r}
#| label: prop-weekend
#| echo: true


prop1 <- round(100 * prop.table(tab1, margin = 1), 1)
rownames(prop1) <- c("Weekday", "Weekend")
colnames(prop1) <- c("No purchase (%)", "Purchase (%)")
prop1
```

The table shows that visitors during the weekend have a higher conversion rate than on weekdays. Notably, while weekend conversion is higher than on weekdays, the difference is only about three percentage points.

## Analysis 2: New vs. Returning Customers

We examine whether returning customers have a better conversion rate than new customers.

```{r}
#| label: analysis-visitor-type
#| echo: true

# Run chi-square test
tab2 <- table(OSPI$VisitorType == "Returning_Visitor", OSPI$Revenue)
testi2 <- chisq.test(tab2)
tulokset2 <- c(testi2$statistic, testi2$p.value)
names(tulokset2) <- c("Chi-square", "p-value")
round(tulokset2, digits = 10)
```

The p-value is essentially 0, so the null hypothesis is rejected and we conclude there is a statistical relationship. Let's review this again in a probability table:

```{r}
#| label: prop-visitor-type
#| echo: true

# Calculate percentages
prop2 <- round(100 * prop.table(tab2, margin = 1), 1)
rownames(prop2) <- c("New customer", "Returning customer")
colnames(prop2) <- c("No purchase (%)", "Purchase (%)")
prop2
```

The conversion rate is clearly higher for new customers. The data do not allow a more detailed look at the reasons.

## Analysis 3: Traffic type and purchase

We examine whether purchasing depends on the channel from which the customer arrived.

```{r}
#| label: analysis-traffic-type
#| echo: true


# Run chi-square test
tab3 <- table(OSPI$TrafficType, OSPI$Revenue)
testi3 <- chisq.test(tab3)

tulokset3 <- c(testi3$statistic, testi3$p.value)
names(tulokset3) <- c("Chi-square", "p-value")
round(tulokset3, digits=5)
```

Some traffic sources are rare, so the expected count in some cells of the contingency table is small. For that reason the χ² test result can be distorted. The large test statistic and small p-value nevertheless suggest there may be a statistical relationship here.

Let's look at the best channels:

```{r}
#| label: conversion-rates
#| echo: true

# Calculate conversion rates
prop3 <- 100 * prop.table(tab3, margin = 1)
conv <- round(prop3[, "TRUE"], 1)
conv_sorted <- sort(conv, decreasing = TRUE)

# Top 5 conversion rates
head(conv_sorted, 5)
```

Traffic type 16 has the best conversion rate, but the number of visitors is small, which makes the conversion unreliable.

### Sales channel summary

Let's create a sales channel summary that accounts for both conversion rate and number of visitors.

```{r}
#| label: traffic-summary
#| echo: true

# Build summary
traffic_summary <- data.frame(
  TrafficType = names(table(OSPI$TrafficType)),
  Visits = as.vector(table(OSPI$TrafficType))
)

purchases <- as.vector(
  tapply(OSPI$Revenue, OSPI$TrafficType, sum)
)

traffic_summary$Purchases <- purchases
traffic_summary$ConversionRate <- round(
  100 * traffic_summary$Purchases / traffic_summary$Visits, 1
)

# Order by purchases
print(
  traffic_summary[order(traffic_summary$Purchases, decreasing = TRUE), ],
  row.names = FALSE
)
```

**Interpretation:** Traffic type 2 appears to be the best option for sales when both conversion rate and visitor count are considered. Looking at the five best conversion rates, traffic type 2 seems like a solid sales channel, especially given that it also has the largest number of visitors.

## Summary

The analysis shows:

1.  **Weekend vs. weekday:** The conversion rate is slightly higher on weekends than on weekdays, but the difference is small (about three percentage points).

2.  **New vs. returning customers:** New customers have a clearly higher conversion rate than returning customers.

3.  **Traffic type:** Traffic type 2 is the best sales channel when both conversion rate and visitor count are considered. Possible additional costs are not considered in this analysis.
